A Framework for Reproducible Data Science using Nix

### 2025-10-20 paper.qmd
citations

Before continuing, it should be noted that other packages exist which provide
similar functionality to [renv]{.pkg}: there is [groundhog]{.pkg} by
@simonsohn2023 which makes it rather easy to install packages as they were on
CRAN at a given date. For example, the code snippet below installs the
[purrr]{.pkg} and [ggplot2]{.pkg} packages as they were on April 4th, 2017:

@@

To illustrate these capabilities with a realistic research task, we present a
polyglot pipeline that simulates a Real Business Cycle (RBC) model in
[Julia]{.proglang}, uses the resulting data to train an [XGBoost]{.pkg}
forecasting model in [Python]{.proglang}, visualizes the results in
[R]{.proglang} with [ggplot2]{.pkg}, and finally compiles a [Quarto]{.pkg}
report.

@@

![Graphical representation of the polyglot pipeline. Green nodes are
[Julia]{.proglang}, yellow nodes are [Python]{.pkg}, and blue nodes are
[R]{.proglang} derivations.](dag-polyglot.png){#fig-dag}

@@

This Julia script implements the state-space solution to the RBC model as a pure
function. The economic theory is based on @depaoli2009.

>>>>

Before continuing, it should be noted that other packages exist which provide
similar functionality to [renv]{.pkg}: there is [groundhog]{.pkg} by
@simonsohn2023 which makes it rather easy to install packages as they were on
CRAN at a given date. For example, the code snippet below installs the
[purrr]{.pkg} [@purrr] and [ggplot2]{.pkg} [@ggplot2] packages as they were on
April 4th, 2017:

@@

To illustrate these capabilities with a realistic research task, we present a
polyglot pipeline that simulates a Real Business Cycle (RBC) model in
[Julia]{.proglang} [@bezanson2017julia], uses the resulting data to train an
[XGBoost]{.pkg} forecasting model in [Python]{.proglang}, visualizes the results
in [R]{.proglang} with [ggplot2]{.pkg}, and finally compiles a [Quarto]{.pkg}
report.

@@

![Graphical representation of the polyglot pipeline. Purple nodes are
[Julia]{.proglang}, yellow nodes are [Python]{.pkg}, light blue nodes are
[R]{.proglang}, and dark blue nodes are Quarto derivations.](dag.png){#fig-dag}

@@

This Julia script implements the state-space solution to the RBC model as a pure
function. The economic theory is based on @depaoli2009rbc.
